---
title: "Description"
author: "Xue Yang"
date: "12/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library("leaps")
```

```{r, message = FALSE}
cancer_df =
  read_csv("Cancer_Registry.csv") %>% 
  janitor::clean_names() %>% 
  dplyr::select(target_death_rate, everything())

raw_data =
  cancer_df %>%
  dplyr::select(-pct_employed16_over, -pct_private_coverage_alone, -binned_inc) %>% 
  dplyr::select(-geography, -avg_deaths_per_year, -pop_est2015, -pct_no_hs18_24 , -pct_hs18_24 , -pct_bach_deg18_24, -pct_some_col18_24, -median_age, -pct_private_coverage, -pct_public_coverage, -pct_public_coverage_alone, -percent_married , -birth_rate) %>% 
  dplyr::select(target_death_rate, everything()) 
  

```


## Descriptive and visualization

**Check for missing value**

```{r}
# calculate the number of missing value and the percentage of the missing value of them
cancer_df %>% 
  summarise_all(funs(sum(is.na(.)))) %>% 
  gather(term, num_na, everything()) %>% 
  mutate(percent = num_na/nrow(cancer_df)) %>% 
  filter(percent !=0) %>% 
  knitr::kable(digits = 3)
```

delete pct_some_col18_24 and pct_employed16_over, pct_private_coverage_alone 

```{r}
cancer_df = 
  cancer_df %>% 
  dplyr::select(-pct_some_col18_24, -pct_employed16_over, -pct_private_coverage_alone)
```

delete geography, binnedlnc

```{r}
cancer_df = 
  cancer_df %>% 
  dplyr::select(-geography, - binned_inc)
```

group the variables

```{r}
# age
age = 
  cancer_df %>% 
  dplyr::select(target_death_rate, median_age, median_age_male, median_age_female)

# edu
edu = 
  cancer_df %>% 
  dplyr::select(target_death_rate, pct_no_hs18_24, pct_hs18_24, pct_bach_deg18_24, pct_hs25_over, pct_bach_deg25_over)

# employ

emp = 
  cancer_df %>% 
  dplyr::select(target_death_rate, pct_unemployed16_over)

# coverage

coverage  =
  cancer_df %>% 
  dplyr::select(target_death_rate, pct_private_coverage, pct_emp_priv_coverage, pct_public_coverage , pct_public_coverage_alone )

# race
race =
  cancer_df %>% 
  dplyr::select(target_death_rate, pct_white, pct_black, pct_asian, pct_other_race)

# marry 
marry = 
  cancer_df %>% 
  dplyr::select(target_death_rate, pct_married_households, percent_married)

# other
other = 
  cancer_df %>% 
  dplyr::select(target_death_rate, avg_ann_count, avg_deaths_per_year, incidence_rate, med_income, pop_est2015, poverty_percent, study_per_cap, avg_household_size, birth_rate)

```


```{r}
cor(age)
pairs(age)

cor(edu)
pairs(edu)

cor(emp)
pairs(emp)

cor(coverage)
pairs(coverage)

cor(race)
pairs(race)

cor(marry)
pairs(marry)

cor(other)
pairs(other)
```

```{r}
cancer_df %>% 
  ggplot(aes(x = target_death_rate)) +
  geom_histogram() +
  labs(
    title = "Histogram of Target Death Rate",
    x = "Target Death Rate",
    y = "Frequency"
  )+
theme_bw() + 
  theme(legend.position = "bottom")
```

From the plot of the distribution of "total cost", we can find that it is significant right skewed, in other words, the distribution of "total cost" is NOT symmetrical at all.

We can do some descriptive statistics for continuous variables:

```{r}
# descriptive statistics for continuous variables
#age
age_list = list(age$median_age, age$median_age_male, age$median_age_female)
min = sapply(age_list, min)
max = sapply(age_list, max)
mean = sapply(age_list, mean)
median = sapply(age_list, median)
sd = sapply(age_list, sd)
IQR = sapply(age_list, IQR)
age_description = rbind(min, max, mean, median, sd, IQR)
colnames(age_description) = c("median_age","median_age_male","median_age_female")
age_description %>% 
  knitr::kable(digits = 3)

# edu 
edu_list = list(edu$pct_no_hs18_24, edu$pct_hs18_24, edu$pct_bach_deg18_24, edu$pct_hs25_over, edu$pct_bach_deg25_over)
min = sapply(edu_list, min)
max = sapply(edu_list, max)
mean = sapply(edu_list, mean)
median = sapply(edu_list, median)
sd = sapply(edu_list, sd)
IQR = sapply(edu_list, IQR)
edu_description = rbind(min, max, mean, median, sd, IQR)
colnames(edu_description) = c("pct_no_hs18_24","pct_hs18_24", "pct_bach_deg18_24", "pct_hs25_over", "pct_bach_deg25_over")
edu_description %>% 
  knitr::kable(digits = 3)

# employ
emp_list = list(emp$pct_unemployed16_over)
min = sapply(emp_list, min)
max = sapply(emp_list, max)
mean = sapply(emp_list, mean)
median = sapply(emp_list, median)
sd = sapply(emp_list, sd)
IQR = sapply(emp_list, IQR)
emp_description = rbind(min, max, mean, median, sd, IQR)
colnames(emp_description) = c("pct_unemployed16_over")
emp_description %>% 
  knitr::kable(digits = 3)

# coverage
coverage_list = list(coverage$pct_private_coverage, coverage$pct_emp_priv_coverage, coverage$pct_public_coverage, coverage$pct_public_coverage_alone)
min = sapply(coverage_list, min)
max = sapply(coverage_list, max)
mean = sapply(coverage_list, mean)
median = sapply(coverage_list, median)
sd = sapply(coverage_list, sd)
IQR = sapply(coverage_list, IQR)
coverage_description = rbind(min, max, mean, median, sd, IQR)
colnames(coverage_description) = c("pct_private_coverage", "pct_emp_priv_coverage", "pct_public_coverage", "pct_public_coverage_alone")
coverage_description %>% 
  knitr::kable(digits = 3)

# race
race_list = list(race$pct_white, race$pct_black, race$pct_asian, race$pct_other_race)
min = sapply(race_list, min)
max = sapply(race_list, max)
mean = sapply(race_list, mean)
median = sapply(race_list, median)
sd = sapply(race_list, sd)
IQR = sapply(race_list, IQR)
race_description = rbind(min, max, mean, median, sd, IQR)
colnames(race_description) = c("pct_white", "pct_black", "pct_asian", "pct_other_race")
race_description %>% 
  knitr::kable(digits = 3)

# marry
marry_list = list(marry$pct_married_households, marry$percent_married)
min = sapply(marry_list, min)
max = sapply(marry_list, max)
mean = sapply(marry_list, mean)
median = sapply(marry_list, median)
sd = sapply(marry_list, sd)
IQR = sapply(marry_list, IQR)
marry_description = rbind(min, max, mean, median, sd, IQR)
colnames(marry_description) = c("pct_married_households","percent_married")
marry_description %>% 
  knitr::kable(digits = 3)

#other
other_list = list(other$avg_ann_count, other$avg_deaths_per_year, other$incidence_rate, other$med_income, other$pop_est2015, other$poverty_percent, other$study_per_cap, other$avg_household_size, other$birth_rate)
min = sapply(other_list, min)
max = sapply(other_list, max)
mean = sapply(other_list, mean)
median = sapply(other_list, median)
sd = sapply(other_list, sd)
IQR = sapply(other_list, IQR)
other_description = rbind(min, max, mean, median, sd, IQR)
colnames(other_description) = c("avg_ann_count", "avg_deaths_per_year", "incidence_rate", "med_income", "pop_est2015", "poverty_percent", "study_per_cap", "avg_household_size", "birth_rate")
other_description %>% 
  knitr::kable(digits = 3)
```

```{r}
age %>% 
  gather(key = variable, value = obervation, -target_death_rate) %>% 
  ggplot(aes(x = variable, y = obervation))+
  geom_boxplot() +
  labs(
    title = "Boxplot for age"
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")

edu %>% 
  gather(key = variable, value = obervation, -target_death_rate) %>% 
  ggplot(aes(x = variable, y = obervation))+
  geom_boxplot() +
  labs(
    title = "Boxplot for edu"
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")

emp %>% 
  gather(key = variable, value = obervation, -target_death_rate) %>% 
  ggplot(aes(x = variable, y = obervation))+
  geom_boxplot() +
  labs(
    title = "Boxplot for employ"
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")

coverage %>% 
  gather(key = variable, value = obervation, -target_death_rate) %>% 
  ggplot(aes(x = variable, y = obervation))+
  geom_boxplot() +
  labs(
    title = "Boxplot for coverage"
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")

race %>% 
  gather(key = variable, value = obervation, -target_death_rate) %>% 
  ggplot(aes(x = variable, y = obervation))+
  geom_boxplot() +
  labs(
    title = "Boxplot for race"
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")

marry %>% 
  gather(key = variable, value = obervation, -target_death_rate) %>% 
  ggplot(aes(x = variable, y = obervation))+
  geom_boxplot() +
  labs(
    title = "Boxplot for marry"
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom")

other %>% 
  gather(key = variable, value = obervation, -target_death_rate) %>% 
  ggplot(aes(x = variable, y = obervation))+
  geom_boxplot() +
  labs(
    title = "Boxplot for other"
  ) + 
  theme_bw() + 
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

```

