---
title: "Model Improvement"
author: "Bihui Sun"
date: "December 15, 2018"
output: github_document
---

```{r}
library(tidyverse)

cancer_df =
  read_csv("Cancer_Registry.csv") %>% 
  janitor::clean_names()

raw_data =
  cancer_df %>%
  dplyr::select(-pct_employed16_over, - pct_private_coverage_alone, -binned_inc) %>% 
  dplyr::select(-geography, -avg_deaths_per_year, -pop_est2015, -pct_no_hs18_24 , -pct_hs18_24 , -pct_bach_deg18_24, -pct_some_col18_24, -median_age, -pct_private_coverage, -pct_public_coverage, -pct_public_coverage_alone, -percent_married , -birth_rate) %>% 
  dplyr::select(target_death_rate, everything()) 

raw_data2 = raw_data %>%
 mutate(pct_other_race_1 = ifelse(pct_other_race > 0.8262, 1, 0))

```

```{r}
fit2 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_other_race + pct_married_households, data = raw_data)
summary(fit2)
fit2_bin = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_other_race_1 + pct_married_households, data = raw_data2)
summary(fit2_bin)
```
* From the previous two rmd we know that we should choose the model 2 according to parsimony. Now we want to try interaction based on model 2. We recode the pct_other_race to 0 and 1 based on median. 0 indicates fewer other race people in this county, and 1 indicates more other race people in this county. We try to interact the new variable pct_other_race_1 to other predictors one by one and see if there is association between race and these predictors.

```{r}
try1 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * avg_ann_count, data = raw_data2)
summary(try1)

```
* First we try race and avg_ann_count. The p-value is 0.153.

```{r}
try2 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * incidence_rate, data = raw_data2)
summary(try2)

```
* The p-value for pct_other_race_1 * incidence_rate is 0.0014, indicating the race modifies the relationship between incidece rate and target mortality rate.

```{r}
try3 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * poverty_percent, data = raw_data2)
summary(try3)

```
* P-value is very small, R square increase a lot.

```{r}
try4 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * pct_hs25_over, data = raw_data2)
summary(try4)
```
* P-value is too big.

```{r}
try5 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * pct_bach_deg25_over, data = raw_data2)
summary(try5)

```
* P-value is small.

```{r}
try6 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * pct_unemployed16_over, data = raw_data2)
summary(try6)

```
* P-value is very small, R square increase a lot.

```{r}
try7 = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * pct_married_households, data = raw_data2)
summary(try7)

```
* P-value is very large.

```{r}
try = lm(target_death_rate ~ avg_ann_count + incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_married_households + pct_other_race_1 + pct_other_race_1 * pct_married_households + pct_other_race_1 * poverty_percent, data = raw_data2)
summary(try)

```
* After adding two most significant interactions, the R squared increase from 0.4853 to 0.4979. 

```{r}
trytry=lm(target_death_rate ~ incidence_rate + poverty_percent + median_age_female + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + pct_other_race + pct_married_households, data = raw_data)
summary(fit2)